AA Distributed Software Development Methodology

### 2013-02-18 paper.tex
foda: AA eh Anarquia Algoritmica

\subsection{Peer Validation}

In order to prevent spamming and to improve the overall quality of AA reports,
each AA session must be validated by another developer. More specifically,
all reports are read by someone that will mark them as `valid' or `invalid'
and may optionally write commentaries about the specific session and quality of
micrologs. The developer in charge of validating any given session is randomly assigned 
by the AA Web server, which sends an email to the chosen developer with an URL to
a validation interface.

@@

In one month, each developer officially contributed to one or many free software
projects. Many developers started the initiation training with no knowledge of what
is free software and ended that period becoming a free software developer.

During that month, the same team developed the first version of the AA
system and used AA to manage their activities. Even while developing
the system. All the source code of AA --- the client that sends the
logs and the AA Web server --- is public available~\footnote{AA source
  code:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
and all the AA sessions log of the whole team of ``Lab Macambira'' is
also on-line~\footnote{Logs of AA sessions:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.

After the training period, during more 6 months, the ``Macambiras''
worked on a large range of free software projects, distributed on work
groups --- each work group focusing on a specific theme like video,
audio and web --- and financed by contracts and support of the
``Pont\~{a}o N\'{o}s Digitais''. In Table~\ref{tabela:criados} we can
see a list of the free software created by ``Lab Macambira'' since
July of 2011.

\begin{table}
    \caption{Software projects created by ``Lab Macambira'' since July
      of 2011 with a short description and the technologies --- like
      programming languages or frameworks --- involved. It is
      interesting to note the heterogeneity of projects and its areas
      of application.}
    \begin{tabular}{|l|p{5cm}|l|}
        \hline
        Application & Description & Technologies involved \\ 
        \hline \hline
        AA            & Algorithmic Auto-regulation      & Python, PHP \\
        \hline
        \'{A}gora Communs & System for on-line deliberations & PHP \\
        \hline
        SIP           & Scilab Image Processing toolbox & C, Scilab \\
        \hline
        animal        & An Imaging Library              & C \\
        \hline
        TeDi          & Test Framework for Distance Transform
        Algorithms & C, Shellscript, Scilab \\
        \hline
        Macambot      & Multi-use IRC Bot               & Python \\
        \hline
        ``Confer\^{e}ncia Permanente'' & Platform for the permanent
        conference of the rights of minors & PHP, JavaScript \\
        \hline
        CPC           & Center for accounting of the Brazilian culture
        representation groups & Python, Django \\
        \hline
        Timeline      & Interactive time lines on the Web & JavaScript
        \\
        \hline
        Imagemap      & Interactive marking for on-line photos &
        JavaScript \\
        \hline
        ABT           & Program for real-time execution and musical
        rhythmic analysis & Python \\
        \hline
        EKP           & Emotional Kernel Panic & Python, ChucK \\
        \hline
        SOS           & Aggregation and diffusion of popular and native
        knowledge about health & Python, Django \\
        \hline
        Creative Economy & Platform for creative, collaborative and
        solidarity economy of the culture hubs and cultural entities &
        Python, Django \\
        \hline
        OpenID Integration & Adaptations to existing software for
        unified login through OpenID & PHP \\
        \hline
        pAAinel & Dashboard for the real-time visualization of Lab
        Macambira activity & Python, Django \\
        \hline
        Georef & Collection of scripts to be used as reference, which
        aims to be a GIS platform to map public data of use to
        citizens & Python, Django \\
        \hline
        AirHackTable & Software for an instrument which generates
        sound from flying origami tracked by webcams & Puredata,
        C/C++, Scilab \\
        \hline
        \end{tabular}
    \label{tabela:criados}
\end{table}

As of this writing the ``Lab Macambira'' have many software
developers, and some of the trained developers continue to work
voluntarily in the project.

@@

%% brief introduction
In a scenario where GCD is growing as a popular form of software
development, not just on free software projects but in the whole
software industry, we need methodologies to deal with its
disadvantages and at the same time to amplify its advantages.

This paper has presented a methodology to GCD, being the development
conducted on large or small groups of software developers, working on
different countries or even at the same room. The AA methodology
implements a simple system where each developer take notes of his work
generating a periodical log of small text sentences. The sum of those
sentences, along an entire session of work, results in a complete
report. The report is made public available through a Website and be
validated by other developers sorted aleatory by the AA Web server.

Instead of a merely work-management tool, AA act as a methodology to
improve the time sense of individuals, dividing their work on small
sessions, and also reducing the need of extensive
reports or unnecessary meetings. By asking users to write a minimal
text sentence as a continuously log, AA does not disturb developers
concentrated on programming: developers just have to type some
characters, hit \textit{enter} and go back to coding.

AA application is not restricted to software development. As of this
writing there is a comic book studio~\footnote{Pula pirata:
  \url{http://pulapirata.com}} starting to use AA to manage their
activities. People with non-software background, like social
scientists, musicians and activists has also using AA and contributing
for its improvement.

For developer teams, we have experienced the use of AA to
auto-regulate the work of ``Lab Macambira'', a group of free software
developers from Brazil. Since July of 2011 the group have contributed
and created new free and open source software for a vast number of
applications.

There are many aspects of the work which remain unfinished. New ways
to report logs --- the ``Twitter like'' messages --- from different
interfaces like IRC, Internet Messaging services and email can make
the use of AA easy and widespread, turning AA an ubiquitous system,
presented on everyday communication channels. Even the work logs
generated since July of 2011 could be vastly statistically analyzed
aiming to recognize patterns in the behavior of individuals and their
productions.

We would like to conclude setting an important role of AA: being a
free software system and an open methodology, AA could be used to
auto-manage groups of individuals working on software or other kinds
of activities. In this way, we are interested to spread AA for those
groups, to have even more developers contributing in a collaborative
way.

@@

We would like to also thank AA: the present research and even this
manuscript was written using AA. The complete log is on-line at
\url{http://www.pulapirata.com/skills/aa}.

>>>>

\subsection{Peer Validation and Peer Pressure}

No set bosses or leaders are required in an ideal aplication the AA methodology;
in practice the role of bosses basically disappears or is greatly lightened --
hence the name Algorithmic \emph{Autoregulation} and other implicit
interpretations to the AA acronym and symbol. The primary mechanism to achieve
AA is peer coordination and social behavior, be it deliberate or implicit. In
order to prevent spamming and to improve the overall quality of AA reports, each
AA session must be validated by another developer. More specifically, all
reports are read by someone that will mark them as `valid' or `invalid' and may
optionally write commentaries about the specific session and quality of
micrologs. The developer in charge of validating any given session is randomly
assigned by the AA Web server, which sends an email to the chosen developer with
an URL to a validation interface.

@@

In one month, each developer officially contributed to one or many free software
projects. Many developers started the initiation training with no knowledge of what
is free software and ended that period becoming a free software developer.
During that month, the same team of trainees also developed the first version of the AA
system and used AA to manage their activities, even while developing
the other aforementioned free software projects.

The source code of AA --- both the client that sends the
logs and the AA Web server --- is public available as free software~\footnote{AA source
  code:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
and all the actual AA log data of the entire team of Lab Macambira from 2011 to
the present time is also on-line~\footnote{Logs of AA sessions:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.

After the initial training period of 1 month, the ``Macambiras''
worked during 6 additional months on a large range of free software projects, distributed on work
groups --- each work group focusing on a specific theme like video,
audio and web --- and funded by contracts, freelance, and support of the
Pont\~{a}o N\'{o}s Digitais~\footnote{\url{http://nosdigitais.teia.org.br}}. 
Table~\ref{tabela:criados} has a list of \emph{new} free software applications created by
``Lab Macambira'' since July 2011.

\begin{table}
    \caption{Software projects created by ``Lab Macambira'' since July
      2011 with a short description and the technologies involved (\eg,
      programming languages or frameworks). It is
      interesting to note the heterogeneity of projects and their areas
      of application.}
    \begin{tabular}{|l|p{5cm}|l|}
        \hline
        Application & Description & Technologies involved \\ 
        \hline \hline
        AA            & Algorithmic Auto-regulation      & Python, PHP \\
        \hline
        \'{A}gora Communs & System for on-line deliberations & PHP \\
        \hline
        SIP           & Scilab Image Processing toolbox & C, Scilab \\
        \hline
        animal        & An Imaging Library              & C \\
        \hline
        TeDi          & Test Framework for Distance Transform
        Algorithms & C, Shellscript, Scilab \\
        \hline
        Macambot      & Multi-use IRC Bot               & Python \\
        \hline
        ``Confer\^{e}ncia Permanente'' & Platform for the permanent
        conference of the rights of minors & PHP, JavaScript \\
        \hline
        CPC           & Center for accounting of the Brazilian culture
        representation groups & Python, Django \\
        \hline
        Timeline      & Interactive time lines on the Web & JavaScript
        \\
        \hline
        Imagemap      & Interactive marking for on-line photos &
        JavaScript \\
        \hline
        ABT           & Program for real-time execution and musical
        rhythmic analysis & Python \\
        \hline
        EKP           & Emotional Kernel Panic & Python, ChucK \\
        \hline
        SOS           & Aggregation and diffusion of popular and native
        knowledge about health & Python, Django \\
        \hline
        Creative Economy & Platform for creative, collaborative and
        solidarity economy of the culture hubs and cultural entities &
        Python, Django \\
        \hline
        OpenID Integration & Adaptations to existing software for
        unified login through OpenID & PHP \\
        \hline
        pAAinel & Dashboard for the real-time visualization of Lab
        Macambira activity & Python, Django \\
        \hline
        Georef & Collection of scripts to be used as reference, which
        aims to be a GIS platform to map public data of use to
        citizens & Python, Django \\
        \hline
        AirHackTable & Software for an instrument which generates
        sound from flying origami tracked by webcams & Puredata,
        C/C++, Scilab \\
        \hline
        \end{tabular}
    \label{tabela:criados}
\end{table}

As of this writing the `Lab Macambira' unites 15 software developers, and 
key 2011-initiated developers continue to work voluntarily in the project.

@@

%% brief introduction
In a scenario where Global Software Development is growing as a popular 
software development scenario both for free software projects and for the entire
software industry, we need methodologies to deal with its potential disadvantages and at
the same time to amplify its advantages.

This paper has presented a methodology for GSD, \ie, the development
involving a series of large or small groups of software developers, each either working on
different countries or at the same room. The AA methodology
implements a simple system where each developer take notes of his work
posting a periodic log of small text sentences or microblogs. The sum of those
micrologs, along with an entire session of work, results in a complete
report. The report is made public available through a Website and be
validated by other developers randomly sorted by the AA Web server.

AA is not limited to a merely work-management tool, but acts as \emph{a
methodology to improve the time sensibility of individuals}, dividing their
complex work on small chunks or sessions, and also reducing the need of extensive
reports or unnecessary meetings. By asking users to write a minimal
text sentence as a continuous log feed, the proposed methodology avoids
disturbing the flow of developers which are heavily concentrated on programming: developers
just have to type a few words and go back to coding.

The AA methodology is not restricted to software development, even though it was
built for the latter. As of this
writing there is an entertainment studio, Pula Pirata~\footnote{
\url{www.pulapirata.com}}, that has started to use AA to manage their creative activities.
Other people with no software background, like social scientists, musicians and
activists also have been using AA and contributing for its improvement.

For developer teams, we have analyzed data and gained practical experience on
the use of AA to auto-regulate the work of Lab Macambira, a group of free
software developers from Brazil. Since July of 2011 the group has contributed
and created new free and open source software for a vast number of additional applications.

There are many aspects of the work which remain unfinished. Other client
interfaces to report micrologs -- the `Twitter-like' messages -- from different
interfaces beyond IRC, \eg, internet social services and email, can make the use of
AA easy and widespread, turning AA an ubiquitous system, presented on everyday
communication channels. Even the actual work logs generated by the Lab Macambira
and Pula Pirata collectives since July of 2011 could be
vastly statistically analyzed aiming to recognize patterns in the behavior of
individuals and their productions.

@@

We would like to also thank AA: the present research and even this
manuscript was written using AA. The complete log is on-line at
\url{www.pulapirata.com/skills/aa}.
